---
title: "Integrated PVA: Streamlined Conservation Portfolio Evaluation"
author: "Anna Ortega (Built on methods from Tomo Eguchi, Summer Martin, Zach Siders)"
date: today
format: 
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-float: true
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  warning: false
  message: false
---

## Overview

Welcome to the **Integrated Population Viability Analysis (PVA)** pipeline. Designed for generic long-lived protected populations (e.g., marine turtles), this 100% self-contained, data-free vignette simulates a multi-threat baseline, implements symmetric lifecycle algebra, evaluates independent conservation options, and exports automated publication-ready graphics.

This pipeline is designed for ultimate reproducibility. It generates its own synthetic historical data, imputes missing values using Bayesian state-space models, and runs forward simulations to evaluate multi-faceted management portfolios.

---

## Phase I: Initialization and Biological Constraints

First, we load our core packages and establish the underlying Conway-Maxwell-Poisson (CMP) mathematical utilities that keep our equations stable. 

```{r setup-and-math}
#| label: setup-and-math
#| code-summary: "Load Packages & Define CMP Math Utilities"

# --- Load core packages ---
library(dplyr)
library(tidyr)
library(ggplot2)
library(mvtnorm)
library(truncnorm)
library(jagsUI)
library(knitr)

# --- Establish directory routing  ---
out_dir <- paste0(getwd(), "/output")
if(!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

# ==============================================================================
# 0. Fixed CMP (Conway-Maxwell-Poisson) mathematical utilities
# ==============================================================================
compute_CMP_constant <- function(Lambda, Nu, Mu, Tol, Max, Log=TRUE, Type="Z"){
  if( (!is.na(Lambda) & Lambda > 10^Nu) | (!is.na(Mu) & Mu^Nu > 10^Nu) ){
    if(Type=="Z"){ ln_Const = Nu*Lambda^(1/Nu) - ((Nu-1)/(2*Nu))*log(Lambda) - ((Nu-1)/2)*log(2*pi) - (1/2)*log(Nu) }
    if(Type=="S"){ ln_Const = Nu*Mu - ((Nu-1)/(2))*log(Mu) - ((Nu-1)/2)*log(2*pi) - (1/2)*log(Nu) }
  }else{
    Const = rep(0,Max+1); Index = 1; Const[Index] = 1
    while( Const[Index]/Const[1] > Tol ){
      if(Type=="Z") Const[Index+1] = Const[Index] * ( Lambda / Index^Nu )
      if(Type=="S") Const[Index+1] = Const[Index] * ( Mu / Index )^Nu; Index = Index + 1
    }
    ln_Const = log(sum(Const))
  }
  if(Log) return(ln_Const) else return(exp(ln_Const))
}

dCMP <- function( x, lambda, mu, nu, log=TRUE, tol=0.01, iter.max=200 ){
  if(missing(mu) & !missing(lambda)) loglike = x*log(lambda) - nu*lfactorial(x) - compute_CMP_constant(Lambda=lambda, Nu=nu, Mu=NA, Tol=tol, Max=iter.max, Log=TRUE, Type="Z")
  if(!missing(mu) & missing(lambda)) loglike = nu*x*log(mu) - nu*lfactorial(x) - compute_CMP_constant(Lambda=NA, Nu=nu, Mu=mu, Tol=tol, Max=iter.max, Log=TRUE, Type="S")
  if(log) return(loglike) else return(exp(loglike))
}

rCMP <- function( n, lambda, mu, nu, tol=0.01, x_max=200 ){
  loglike_x = rep(NA, x_max+1)
  for( x in 0:x_max ){
    if(missing(mu) & !missing(lambda)) loglike_x[x+1] = dCMP( x=x, lambda=lambda, nu=nu, log=TRUE, tol=tol, iter.max=x_max)
    if(!missing(mu) & missing(lambda)) loglike_x[x+1] = dCMP( x=x, mu=mu, nu=nu, log=TRUE, tol=tol, iter.max=x_max)
  }
  return(sample(x=0:x_max, size=n, replace=TRUE, prob=exp(loglike_x)))
}

# --- PRIMARY COHORT DATA MATH TRANSFORMS ---
inv_logit  <- function(x) { exp(x) / (1 + exp(x)) }
safe_logit <- function(x) { x_safe <- pmax(0.001, pmin(0.999, x)); log(x_safe / (1 - x_safe)) }
```

### Defining the Species and Threats
Before projecting into the future, we define the life-history parameters and the conservation interventions we plan to test. 

```{r biology-constants}
#| label: biology-constants
#| code-fold: false

set.seed(42) # Lock randomizer for identical results every run
n_sims        <- 500   # Monte Carlo simulation sample size
horizon       <- 100   # Evaluation horizon window (years)
hist_years    <- 2000:2020
n_hist_yrs    <- length(hist_years)

# --- Standard species demographic rates ---
sp_clutch_freq <- 4.0    # Average nests per female per year
sp_remig_int   <- 3.0    # Adult female remigration interval (years)
sp_ane_pj      <- 0.82   # Juvenile age-class annual survival
sp_ane_pa      <- 0.92   # Mature adult female annual survival
sp_linf        <- 105.0  # Asymptotic maximum Straight Carapace Length (SCL, cm)
sp_k           <- 0.12   # VBGF somatic growth parameter
sp_tknot       <- -1.5   # Theoretical age at length zero (t_0)
sp_lmat        <- 95.0   # Length constraint proxy triggering maturity transition
sp_sig_mat     <- 4.5    # Logistic slope of maturation ogive curve
sp_max_age     <- 50.0   # Asymptotic maximum age lifespan ceiling
sp_pf          <- 0.68   # Sex ratio split (percent female)

# --- Threat parameter constants (Fishery baseline) ---
cmp_mu         <- 15.0   # Status quo mean annual longline interaction rate
cmp_nu         <- 0.15   # CMP dispersion factor

# --- Multi-option conservation portfolio parameters ---
optA_nests_saved       <- 35    # Option A: Extra nests saved annually
sp_clutch_size         <- 95    # Mean eggs deposited per clutch
sp_hatch_success       <- 0.72  # Hatching success probability (under protection)
sp_hatch_to_yearling   <- 0.025 # First-year pelagic survival rate baseline

optB_headstarts_released <- 150 # Option B: Yearling juveniles reared/released
optB_release_age         <- 2.0 # Age at release

optC_adults_saved        <- 6   # Option C: Mature nesting females rescued

optD_cmp_mu_mitigated    <- 4.5 # Option D: Mitigated interaction rate 

optE_mort_min            <- 0.05 # Option E: Lower post-release mortality bounds
optE_mort_max            <- 0.18 # Upper post-release mortality bounds
```

---

## Phase II: Virtual Reality (Mock Data & Imputation)

To test our PVA, we generate synthetic historical data across three monitoring beaches (Alpha, Beta, Gamma). Real-world data is messy, so we purposefully inject a 30% missing data rate into our seasonal wave, and then use `jagsUI` to impute those missing values via a Bayesian state-space model.

```{r mock-data-and-jags}
#| label: mock-data-and-jags
#| code-summary: "Generate Synthetic Data & Run JAGS Imputation"
#| results: hide

# Simulate fishery observer length and mortality measurements
mock_observer <- data.frame(
  Year = sort(sample(hist_years[hist_years >= 2005], 50, replace = TRUE)),
  Species = "Mock_Sea_Turtle",
  SCL_raw = round(rnorm(50, mean = 76.5, sd = 9.8), 1),
  M_mu = round(runif(50, min = 0.25, max = 0.75), 3)
)

# Simulate total fleet-wide interaction estimates (SAFE compliance accounts)
mock_safe <- data.frame(Year = hist_years) %>%
  mutate(Total_Est = ifelse(Year < 2005, 0, round(rnorm(n(), mean = 22, sd = 3))))

# 1. Create a monthly dataset with a simulated seasonal wave
months_total <- n_hist_yrs * 12
mock_monthly <- expand.grid(
  Month_Idx = 1:months_total,
  Site = c("Alpha_Beach", "Beta_Beach", "Gamma_Beach")
) %>%
  mutate(
    Year = min(hist_years) + floor((Month_Idx - 1) / 12),
    Month = (Month_Idx - 1) %% 12 + 1,
    True_Base = ifelse(Site == "Alpha_Beach", 450,
                       ifelse(Site == "Beta_Beach", 280, 180)) * exp(-0.025 * (Year - min(hist_years))),
    Seasonal_Multiplier = pmax(0.1, 1 + 0.8 * cos(2 * pi * (Month - 6.5) / 12)),
    True_Nests = (True_Base / 12) * Seasonal_Multiplier * runif(n(), 0.8, 1.2),
    Observed_Nests = ifelse(runif(n()) < 0.30, NA, round(True_Nests)) # 30% missing data
  )

obs_matrix <- mock_monthly %>%
  select(Month_Idx, Site, Observed_Nests) %>%
  pivot_wider(names_from = Month_Idx, values_from = Observed_Nests) %>%
  select(-Site) %>%
  as.matrix()

# 2. Define the JAGS state-space model with seasonal harmonics
jags_seasonal_model <- "
model {
  for (i in 1:n_sites) {
    X[i, 1] ~ dnorm(mu0, tau0)
    beta_cos[i] ~ dnorm(0, 0.01)
    beta_sin[i] ~ dnorm(0, 0.01)

    for (t in 1:n_months) {
      seasonality[i, t] <- beta_cos[i] * cos(2 * pi * month_seq[t] / period) +
                           beta_sin[i] * sin(2 * pi * month_seq[t] / period)
      mu_obs[i, t] <- X[i, t] + seasonality[i, t]
      Y[i, t] ~ dnorm(mu_obs[i, t], tau_obs)
    }
    for (t in 2:n_months) {
      X[i, t] ~ dnorm(X[i, t-1] + u, tau_proc)
    }
  }
  mu0 ~ dnorm(0, 0.001)
  tau0 <- pow(sigma0, -2)
  sigma0 ~ dunif(0, 100)

  u ~ dnorm(0, 0.001)
  tau_proc <- pow(sigma_proc, -2)
  sigma_proc ~ dunif(0, 100)

  tau_obs <- pow(sigma_obs, -2)
  sigma_obs ~ dunif(0, 100)
}
"

jags_data <- list(
  Y = log(obs_matrix + 1),
  n_sites = nrow(obs_matrix),
  n_months = ncol(obs_matrix),
  month_seq = rep(1:12, times = n_hist_yrs),
  period = 12, pi = pi
)

jags_params <- c("mu_obs", "beta_cos", "beta_sin", "u")
model_file_path <- tempfile(fileext = ".txt")
writeLines(jags_seasonal_model, con = model_file_path)

out_jags <- jags(
  data = jags_data,
  parameters.to.save = jags_params,
  model.file = model_file_path,
  n.chains = 3, n.iter = 5000, n.burnin = 1000, n.thin = 2,
  parallel = TRUE, verbose = FALSE
)
unlink(model_file_path)

# Extract and merge imputed values
imputed_log_matrix <- out_jags$q50$mu_obs
imputed_nests <- exp(imputed_log_matrix) - 1

mock_abundance <- mock_monthly %>%
  arrange(Site, Month_Idx) %>%
  mutate(
    Imputed_Nests = pmax(0, as.vector(t(imputed_nests))),
    Final_Monthly_Nests = ifelse(is.na(Observed_Nests), Imputed_Nests, Observed_Nests)
  ) %>%
  group_by(Year, Site) %>%
  summarise(Annual_Nests = sum(Final_Monthly_Nests), .groups = "drop")
```

With the imputation complete, we extract the background population growth rate (`U`) and process variance (`Q`), generating our baseline trend over the three mock beaches.

```{r trend-plot}
#| label: trend-plot
#| code-fold: false
#| fig-cap: "Manuscript Fig 1: Baseline Trend Fit across simulated nesting beaches."

mock_posteriors <- list(
  U = rnorm(n_sims, mean = -0.028, sd = 0.004),
  Q = rgamma(n_sims, shape = 2, rate = 45) 
)
start_pop_draws <- rnorm(n_sims, mean = 720, sd = 45) * sp_remig_int

p_fit <- ggplot(mock_abundance, aes(x = Year, y = log(Annual_Nests/sp_clutch_freq), color = Site)) +
  geom_point(size = 2) + geom_line(alpha = 0.5) +
  theme_minimal() + labs(y = "log(Annual Nesting Cohort)")

print(p_fit)
ggsave(filename = paste0(out_dir, "/Manuscript_Fig1_Trend_Fit.png"), plot = p_fit, width = 7, height = 4.5)
```

---

## Phase III: The Adult Nester Equivalent (ANE) Calculator

To accurately project population impacts, we have to track the "ghost life" of bycatch. By mapping carapace lengths to an explicit age class, we can calculate what an individual turtle's life *would have looked like* had it not interacted with the fishery, capturing the fractional loss of future reproductive value.

```{r ane-calculator}
#| label: ane-calculator
#| code-summary: "Retrospective Demographic Conversions (ANE)"

all_turtles <- data.frame()
for(yr in unique(mock_safe$Year)) {
  obs_yr    <- mock_observer %>% filter(Year == yr)
  total_est <- mock_safe %>% filter(Year == yr) %>% pull(Total_Est)
  n_unobs   <- max(0, total_est - nrow(obs_yr))

  if(nrow(obs_yr) > 0) {
    all_turtles <- rbind(all_turtles, data.frame(Year = yr, Length = obs_yr$SCL_raw, Mortality = obs_yr$M_mu))
  }
  if(n_unobs > 0) {
    all_turtles <- rbind(all_turtles, data.frame(Year = yr, Length = rnorm(n_unobs, mean = 74, sd = 7), Mortality = runif(n_unobs, 0.45, 0.55)))
  }
}

sp_ledger <- data.frame()
for (i in 1:nrow(all_turtles)) {
  c_year <- all_turtles$Year[i]

  age_start <- sp_tknot - (1 / sp_k) * log(1 - (pmin(all_turtles$Length[i], sp_linf - 0.1) / sp_linf))
  if(is.nan(age_start) || is.na(age_start)) age_start <- sp_max_age - 2

  future_years <- seq(c_year, max(hist_years))
  l_y          <- length(future_years)
  if(l_y < 1) next

  ages_traj <- seq(age_start, length.out = l_y, by = 1)
  lens_traj <- ifelse(ages_traj >= sp_max_age, rep(sp_linf * 0.99, l_y), sp_linf * (1 - exp(-sp_k * (ages_traj - sp_tknot))))

  p_mat   <- 1.0 / (1.0 + exp(-(lens_traj - sp_lmat) / sp_sig_mat))
  p_binom <- rbinom(l_y, 1, p_mat)
  if(any(p_binom == 1)) p_binom[min(which(p_binom == 1)):l_y] <- 1

  surv_vector <- cumprod((1 - p_binom) * sp_ane_pj + p_binom * sp_ane_pa)
  sr <- surv_vector * p_binom * sp_pf * all_turtles$Mortality[i] * (1 / sp_remig_int)
  
  sp_ledger <- rbind(sp_ledger, data.frame(CalendarYear = future_years, ANE = sr))
}

final_historical_ane <- sp_ledger %>% group_by(CalendarYear) %>% summarise(Total_ANE = sum(ANE), .groups = "drop")

mock_posteriors_adj <- list(
  U = mock_posteriors$U + 0.0022, 
  Q = mock_posteriors$Q * 0.94    
)
start_pop_draws_adj <- start_pop_draws + (tail(final_historical_ane$Total_ANE, 1) * sp_remig_int)
```

---

## Phase IV: The 8-Scenario Simulator

We simulate forward for `r horizon` years across 8 independent arrays, isolating specific intervention impacts alongside a combined portfolio strategy.

* **Scenario 1:** Unimpacted pristine baseline
* **Scenario 2:** Status Quo (Current threat, fishery take)
* **Scenarios 3-7:** Independent interventions (Options A - E)
* **Scenario 8:** Combined management portfolio

```{r simulator-loop}
#| label: simulator-loop
#| code-summary: "100-Year Forward Simulation Loop (8 Tracks)"
#| cache: true

sim_notake   <- matrix(NA, n_sims, horizon)
sim_take     <- matrix(NA, n_sims, horizon)
sim_optA     <- matrix(NA, n_sims, horizon)
sim_optB     <- matrix(NA, n_sims, horizon)
sim_optC     <- matrix(NA, n_sims, horizon)
sim_optD     <- matrix(NA, n_sims, horizon)
sim_optE     <- matrix(NA, n_sims, horizon)
sim_combined <- matrix(NA, n_sims, horizon)

for(i in 1:n_sims) {
  curr_nt <- start_pop_draws_adj[i]
  curr_t  <- curr_A <- curr_B <- curr_C <- curr_D <- curr_E <- curr_comb <- start_pop_draws[i]

  vec_loss_sq        <- rep(0, horizon)
  vec_loss_optE      <- rep(0, horizon)
  vec_gain_optA      <- rep(0, horizon)
  vec_gain_optB      <- rep(0, horizon)
  vec_gain_optC      <- rep(0, horizon)

  yearlings_produced_optA <- optA_nests_saved * sp_clutch_size * sp_hatch_success * sp_hatch_to_yearling

  for(y in 1:horizon) {
    curr_nt <- rnorm(1, curr_nt * exp(mock_posteriors_adj$U[i]), sqrt(mock_posteriors_adj$Q[i]))
    sim_notake[i,y] <- curr_nt

    # Status Quo Losses
    takes_sq <- rCMP(1, mu = cmp_mu, nu = cmp_nu)
    if (takes_sq > 0) {
      for(t in 1:takes_sq) {
        future_len <- rnorm(1, mean = 75, sd = 8) 
        age_t <- sp_tknot - (1 / sp_k) * log(1 - (pmin(future_len, sp_linf - 0.1) / sp_linf)) 
        if(is.nan(age_t) || is.na(age_t)) age_t <- sp_max_age - 2
        rem_yrs   <- y:horizon; l_y <- length(rem_yrs)
        ages_traj <- seq(age_t, length.out = l_y, by = 1)
        lens_traj <- ifelse(ages_traj >= sp_max_age, rep(sp_linf * 0.99, l_y), sp_linf * (1 - exp(-sp_k * (ages_traj - sp_tknot))))
        p_mat     <- 1.0 / (1.0 + exp(-(lens_traj - sp_lmat) / sp_sig_mat)) 
        p_binom   <- rbinom(l_y, 1, p_mat)
        if(any(p_binom == 1)) p_binom[min(which(p_binom == 1)):l_y] <- 1

        surv_vector <- cumprod((1 - p_binom) * sp_ane_pj + p_binom * sp_ane_pa) 
        future_mort <- runif(1, 0.3, 0.7)
        sr          <- surv_vector * p_binom * sp_pf * future_mort * (1 / sp_remig_int)
        vec_loss_sq[y:horizon] <- vec_loss_sq[y:horizon] + sr 
      }
    }

    # Option D (Gear avoidance)
    takes_optD <- rCMP(1, mu = optD_cmp_mu_mitigated, nu = cmp_nu)
    vec_loss_optD_annual <- 0
    if (takes_optD > 0) {
      vec_loss_optD_temp <- rep(0, horizon)
      for(t in 1:takes_optD) {
        future_len <- rnorm(1, mean = 75, sd = 8)
        age_t <- sp_tknot - (1 / sp_k) * log(1 - (pmin(future_len, sp_linf - 0.1) / sp_linf))
        if(is.nan(age_t) || is.na(age_t)) age_t <- sp_max_age - 2
        rem_yrs   <- y:horizon; l_y <- length(rem_yrs)
        ages_traj <- seq(age_t, length.out = l_y, by = 1)
        lens_traj <- ifelse(ages_traj >= sp_max_age, rep(sp_linf * 0.99, l_y), sp_linf * (1 - exp(-sp_k * (ages_traj - sp_tknot))))
        p_mat     <- 1.0 / (1.0 + exp(-(lens_traj - sp_lmat) / sp_sig_mat))
        p_binom   <- rbinom(l_y, 1, p_mat)
        if(any(p_binom == 1)) p_binom[min(which(p_binom == 1)):l_y] <- 1

        surv_vector <- cumprod((1 - p_binom) * sp_ane_pj + p_binom * sp_ane_pa)
        future_mort <- runif(1, 0.3, 0.7)
        sr          <- surv_vector * p_binom * sp_pf * future_mort * (1 / sp_remig_int)
        vec_loss_optD_temp[y:horizon] <- vec_loss_optD_temp[y:horizon] + sr
      }
      vec_loss_optD_annual <- vec_loss_optD_temp[y]
    }

    # Option E (Safe handling)
    if (takes_sq > 0) {
      for(t in 1:takes_sq) {
        future_len <- rnorm(1, mean = 75, sd = 8)
        age_t <- sp_tknot - (1 / sp_k) * log(1 - (pmin(future_len, sp_linf - 0.1) / sp_linf))
        if(is.nan(age_t) || is.na(age_t)) age_t <- sp_max_age - 2
        rem_yrs   <- y:horizon; l_y <- length(rem_yrs)
        ages_traj <- seq(age_t, length.out = l_y, by = 1)
        lens_traj <- ifelse(ages_traj >= sp_max_age, rep(sp_linf * 0.99, l_y), sp_linf * (1 - exp(-sp_k * (ages_traj - sp_tknot))))
        p_mat     <- 1.0 / (1.0 + exp(-(lens_traj - sp_lmat) / sp_sig_mat))
        p_binom   <- rbinom(l_y, 1, p_mat)
        if(any(p_binom == 1)) p_binom[min(which(p_binom == 1)):l_y] <- 1

        surv_vector    <- cumprod((1 - p_binom) * sp_ane_pj + p_binom * sp_ane_pa)
        future_mort_me <- runif(1, optE_mort_min, optE_mort_max) 
        sr             <- surv_vector * p_binom * sp_pf * future_mort_me * (1 / sp_remig_int)
        vec_loss_optE[y:horizon] <- vec_loss_optE[y:horizon] + sr
      }
    }

    # Option A Gains
    if (yearlings_produced_optA > 0) { 
      rem_yrs_A <- y:horizon; l_y_A <- length(rem_yrs_A)
      ages_A    <- seq(1.0, length.out = l_y_A, by = 1)
      lens_A    <- ifelse(ages_A >= sp_max_age, rep(sp_linf * 0.99, l_y_A), sp_linf * (1 - exp(-sp_k * (ages_A - sp_tknot))))
      p_mat_A   <- 1.0 / (1.0 + exp(-(lens_A - sp_lmat) / sp_sig_mat))
      p_binom_A <- rbinom(l_y_A, 1, p_mat_A)
      if(any(p_binom_A == 1)) p_binom_A[min(which(p_binom_A == 1)):l_y_A] <- 1
      surv_vector_A <- cumprod((1 - p_binom_A) * sp_ane_pj + p_binom_A * sp_ane_pa)
      vec_gain_optA[y:horizon] <- vec_gain_optA[y:horizon] + (surv_vector_A * p_binom_A * sp_pf * (1 / sp_remig_int) * yearlings_produced_optA)
    }

    # Option B Gains
    if (optB_headstarts_released > 0) {
      rem_yrs_B <- y:horizon; l_y_B <- length(rem_yrs_B)
      ages_B    <- seq(optB_release_age, length.out = l_y_B, by = 1)
      lens_B    <- ifelse(ages_B >= sp_max_age, rep(sp_linf * 0.99, l_y_B), sp_linf * (1 - exp(-sp_k * (ages_B - sp_tknot))))
      p_mat_B   <- 1.0 / (1.0 + exp(-(lens_B - sp_lmat) / sp_sig_mat))
      p_binom_B <- rbinom(l_y_B, 1, p_mat_B)
      if(any(p_binom_B == 1)) p_binom_B[min(which(p_binom_B == 1)):l_y_B] <- 1
      surv_vector_B <- cumprod((1 - p_binom_B) * sp_ane_pj + p_binom_B * sp_ane_pa)
      vec_gain_optB[y:horizon] <- vec_gain_optB[y:horizon] + (surv_vector_B * p_binom_B * sp_pf * (1 / sp_remig_int) * optB_headstarts_released)
    }

    # Option C Gains
    if (optC_adults_saved > 0) {
      rem_yrs_C <- y:horizon; l_y_C <- length(rem_yrs_C)
      ages_C    <- seq(35.0, length.out = l_y_C, by = 1)
      lens_C    <- rep(sp_linf * 0.99, l_y_C)
      p_binom_C <- rep(1, l_y_C)
      surv_vector_C <- cumprod(p_binom_C * sp_ane_pa)
      vec_gain_optC[y:horizon] <- vec_gain_optC[y:horizon] + (surv_vector_C * p_binom_C * (1 / sp_remig_int) * optC_adults_saved)
    }

    # -----------------------------
    # Applying the Math to Tracks
    # -----------------------------
    curr_t <- rnorm(1, (curr_t - vec_loss_sq[y]) * exp(mock_posteriors$U[i]), sqrt(mock_posteriors$Q[i]))
    sim_take[i,y] <- curr_t

    curr_A <- rnorm(1, (curr_A - vec_loss_sq[y] + vec_gain_optA[y]) * exp(mock_posteriors$U[i]), sqrt(mock_posteriors$Q[i]))
    sim_optA[i,y] <- curr_A

    curr_B <- rnorm(1, (curr_B - vec_loss_sq[y] + vec_gain_optB[y]) * exp(mock_posteriors$U[i]), sqrt(mock_posteriors$Q[i]))
    sim_optB[i,y] <- curr_B

    curr_C <- rnorm(1, (curr_C - vec_loss_sq[y] + vec_gain_optC[y]) * exp(mock_posteriors$U[i]), sqrt(mock_posteriors$Q[i]))
    sim_optC[i,y] <- curr_C

    curr_D <- rnorm(1, (curr_D - vec_loss_optD_annual) * exp(mock_posteriors$U[i]), sqrt(mock_posteriors$Q[i]))
    sim_optD[i,y] <- curr_D

    curr_E <- rnorm(1, (curr_E - vec_loss_optE[y]) * exp(mock_posteriors$U[i]), sqrt(mock_posteriors$Q[i]))
    sim_optE[i,y] <- curr_E

    curr_comb <- rnorm(1, (curr_comb - vec_loss_optE[y] + vec_gain_optA[y] + vec_gain_optB[y] + vec_gain_optC[y]) * exp(mock_posteriors$U[i]), sqrt(mock_posteriors$Q[i]))
    sim_combined[i,y] <- curr_comb
  }

  # Absorbing Barriers (extirpation limits)
  if(any(sim_notake[i,] < 0))   { sim_notake[i, min(which(sim_notake[i,] < 0)):horizon]   <- 0 }
  if(any(sim_take[i,] < 0))     { sim_take[i, min(which(sim_take[i,] < 0)):horizon]     <- 0 }
  if(any(sim_optA[i,] < 0))     { sim_optA[i, min(which(sim_optA[i,] < 0)):horizon]     <- 0 }
  if(any(sim_optB[i,] < 0))     { sim_optB[i, min(which(sim_optB[i,] < 0)):horizon]     <- 0 }
  if(any(sim_optC[i,] < 0))     { sim_optC[i, min(which(sim_optC[i,] < 0)):horizon]     <- 0 }
  if(any(sim_optD[i,] < 0))     { sim_optD[i, min(which(sim_optD[i,] < 0)):horizon]     <- 0 }
  if(any(sim_optE[i,] < 0))     { sim_optE[i, min(which(sim_optE[i,] < 0)):horizon]     <- 0 }
  if(any(sim_combined[i,] < 0)) { sim_combined[i, min(which(sim_combined[i,] < 0)):horizon] <- 0 }
}
```

---

## Phase V: Outcomes and Strategic Portfolios

Finally, we aggregate our massive multi-track simulation to see which conservation strategies hold water over the 100-year timeline.

```{r results-compilation}
#| label: results-compilation
#| code-fold: false

yrs_proj <- (max(hist_years) + 1):(max(hist_years) + horizon)

generate_quantiles <- function(mat, label) {
  data.frame(
    Year = yrs_proj, Median = apply(mat, 2, median),
    L95 = apply(mat, 2, function(x) quantile(x, 0.025)),
    U95 = apply(mat, 2, function(x) quantile(x, 0.975)), Scenario = label
  )
}

plot_df <- bind_rows(
  generate_quantiles(sim_notake, "Pristine Baseline (No Take)"),
  generate_quantiles(sim_take, "Status Quo Threats Only"),
  generate_quantiles(sim_optA, "Option A: Nest protection"),
  generate_quantiles(sim_optB, "Option B: Headstarting"),
  generate_quantiles(sim_optC, "Option C: Poaching cessation"),
  generate_quantiles(sim_combined, "Combined Recovery Portfolio")
)
```

### Visualizing the Projections

Here we look at the overlaid outcomes for the individual strategies compared to the powerful combined portfolio option.

```{r plots}
#| label: plots
#| code-fold: false
#| fig-width: 8
#| fig-height: 6

p_overlay <- ggplot(plot_df, aes(x = Year, y = Median, color = Scenario, fill = Scenario)) +
  geom_line(linewidth = 1.2) + geom_ribbon(aes(ymin = L95, ymax = U95), alpha = 0.08, color = NA) +
  theme_bw() + theme(legend.position = "bottom", legend.direction = "vertical") +
  labs(title = "Multi-Option Conservation Portfolio Projections", y = "Total Mature Population Size")

print(p_overlay)
ggsave(filename = paste0(out_dir, "/Manuscript_Fig3_PVA_Portfolio_Overlay.png"), plot = p_overlay, width = 8, height = 6)

df_net_comb <- generate_quantiles(sim_combined - sim_take, "Net Portfolio Gain")
p_diff <- ggplot(df_net_comb, aes(x = Year, y = Median)) +
  geom_line(color = "darkorchid4", linewidth = 1.2) +
  geom_ribbon(aes(ymin = L95, ymax = U95), fill = "darkorchid", alpha = 0.20) +
  theme_bw() + labs(title = "Net Combined Portfolio Payoff Timeline", y = "Net Abundance Gain (\u0394 Population)")

print(p_diff)
ggsave(filename = paste0(out_dir, "/Manuscript_Fig4_Net_Portfolio_Payoff.png"), plot = p_diff, width = 8, height = 4)
```

### 100-Year Scenario Scorecard

At the end of our evaluation horizon, here is how each strategy mathematically performs compared against the others.

```{r scorecard}
#| label: scorecard
#| code-fold: false

scorecard_table <- data.frame(
  Management_Strategy_Track = c(
    "1. Pristine Baseline (No Take)", "2. Status Quo Realized Threats",
    "3. Option A (Nest Protections)", "4. Option B (Headstarting Enhancements)",
    "5. Option C (Poaching Cessation)", "6. Option D (Gear Avoidance Closures)",
    "7. Option E (Handling Severity Mitigation)", "8. Combined Strategic Portfolio"
  ),
  Median_100_Year_Abundance = c(
    median(sim_notake[, horizon]), median(sim_take[, horizon]),
    median(sim_optA[, horizon]), median(sim_optB[, horizon]),
    median(sim_optC[, horizon]), median(sim_optD[, horizon]),
    median(sim_optE[, horizon]), median(sim_combined[, horizon])
  )
)

knitr::kable(scorecard_table, caption = "Table 1: Comprehensive PVA 100-Year Scorecard", digits = 1)
write.csv(scorecard_table, paste0(out_dir, "/Manuscript_Table1_Comprehensive_PVA_Scorecard.csv"), row.names = FALSE)
```